Diel activity GAMs

### 2021-09-28 diel_activity_analysis.Rmd
diel strength % diff

### Spatial variation in the strength of diel patterns

@@

# predict model results into dataframe
data_g_plot <- cbind(data_g_plot, predict.gam(gam_fox_sp, newdata = data_g_plot, se.fit = TRUE, type = "link", exclude = c("s(station)", "s(survey_duration)")))
data_g_plot <- rename(data_g_plot, fox_predicted = fit,  fox_predicted_se = se.fit) # rename
# give each row / location a unique name to group by
data_g_plot$location_id <- paste0(data_g_plot$x, "_", data_g_plot$y)
# get difference in min max diel activity for each row
data_g_plot <- data_g_plot %>% 
  group_by(location_id) %>% 
  mutate(diel_diff = max(fox_predicted) - min(fox_predicted))

# fox plot: spatial activity per hour
g_fox_plot_hr <- ggplot(aes(x, y, fill = fox_predicted),
                     data = data_g_plot) +
  geom_tile()+
  scale_fill_viridis("log(activity)", option = "viridis", limits = c(-5.27, -1.13)) +
  geom_point(data = records_g, fill = NA, col = "white", size = 0.7, alpha = 0.1, shape = 3) +
  theme_bw(10) + 
  theme(axis.title = element_blank()) + 
    labs(title = 'Glenelg region; red fox; Hour: {frame_time}:00') +
  transition_time(hour) +
  ease_aes('linear')

# fox plot: spatial difference
g_fox_plot_diff <- ggplot(aes(x, y, fill = diel_diff),
                     data = data_g_plot) +
  geom_tile()+
  scale_fill_viridis("Relative diel pattern strength", option = "viridis", limits = c(0.5, 1.55), breaks=c(0.5, 1.03, 1.55),labels=c("low", "medium", "high")) +
  geom_point(data = records_g, fill = NA, col = "white", size = 0.7, alpha = 0.015, shape = 3) +
  theme_bw(10) + 
  ggtitle("a.  Glenelg region; red fox") +
  theme(axis.title = element_blank())

@@

# subset to just locations within buffer zone
data_o_plot = data_o_plot[with(data_o_plot, inSide(otways_buffer_df, x, y)),]
data_o_plot <- cbind(data_o_plot, predict.gam(gam_fox_sp, newdata = data_o_plot, se.fit = TRUE, type = "link", exclude = c("s(station)", "s(survey_duration)")))
data_o_plot <- rename(data_o_plot, fox_predicted = fit,  fox_predicted_se = se.fit) # rename
# give each row / location a unique name to group by
data_o_plot$location_id <- paste0(data_o_plot$x, "_", data_o_plot$y)
# get difference in min max diel activity for each row
data_o_plot <- data_o_plot %>% 
  group_by(location_id) %>% 
  mutate(diel_diff = max(fox_predicted) - min(fox_predicted))

# fox plot: spatial activity per hour
o_fox_plot_hr <- ggplot(aes(x, y, fill = fox_predicted),
                     data = data_o_plot) +
  geom_tile()+
  scale_fill_viridis("log(activity)", option = "viridis", limits = c(-5.27, -1.13)) +
  geom_point(data = records_o, fill = NA, col = "white", size = 0.7, alpha = 0.1, shape = 3) +
  theme_bw(10) + 
  theme(axis.title = element_blank()) + 
    labs(title = 'Otway Ranges; red fox; Hour: {frame_time}:00') +
  transition_time(hour) +
  ease_aes('linear')

# fox plot: spatial difference
o_fox_plot_diff <- ggplot(aes(x, y, fill = diel_diff),
                     data = data_o_plot) +
  geom_tile()+
  scale_fill_viridis("Relative diel pattern strength", option = "viridis", limits = c(0.5, 1.55), breaks=c(0.5, 1.03, 1.55),labels=c("low", "medium", "high")) +
  geom_point(data = records_o, fill = NA, col = "white", size = 0.7, alpha = 0.015, shape = 3) +
  theme_bw(10) + 
  ggtitle("c.  Otway Ranges; red fox") +
  theme(axis.title = element_blank())

#### CAT 
### GLENELG
# predict model results into dataframe
data_g_plot <- cbind(data_g_plot, predict.gam(gam_cat_sp, newdata = data_g_plot, se.fit = TRUE, type = "link", exclude = c("s(station)", "s(survey_duration)")))
data_g_plot <- rename(data_g_plot, cat_predicted = fit,  cat_predicted_se = se.fit) # rename
# give each row / location a unique name to group by
data_g_plot$location_id <- paste0(data_g_plot$x, "_", data_g_plot$y)
# get difference in min max diel activity for each row
data_g_plot <- data_g_plot %>% 
  group_by(location_id) %>% 
  mutate(diel_diff = max(cat_predicted) - min(cat_predicted))

# cat plot: spatial activity per hour
g_cat_plot_hr <- ggplot(aes(x, y, fill = cat_predicted),
                     data = data_g_plot) +
  geom_tile()+
  scale_fill_viridis("log(activity)", option = "viridis", limits = c(-5.27, -1.13)) +
  geom_point(data = records_g, fill = NA, col = "white", size = 0.7, alpha = 0.1, shape = 3) +
  theme_bw(10) + 
  theme(axis.title = element_blank()) + 
    labs(title = 'Glenelg region; feral cat; Hour: {frame_time}:00') +
  transition_time(hour) +
  ease_aes('linear')

# cat plot: spatial difference
g_cat_plot_diff <- ggplot(aes(x, y, fill = diel_diff),
                     data = data_g_plot) +
  geom_tile()+
  scale_fill_viridis("Relative diel pattern strength", option = "viridis", limits = c(0.5, 1.55), breaks=c(0.5, 1.03, 1.55),labels=c("low", "medium", "high")) +
  geom_point(data = records_g, fill = NA, col = "white", size = 0.7, alpha = 0.015, shape = 3) +
  theme_bw(10) + 
  ggtitle("b.  Glenelg region; feral cat") +
   theme(axis.title = element_blank())

### OTWAYS
# predict model results into dataframe
data_o_plot <- cbind(data_o_plot, predict.gam(gam_cat_sp, newdata = data_o_plot, se.fit = TRUE, type = "link", exclude = c("s(station)", "s(survey_duration)")))
data_o_plot <- rename(data_o_plot, cat_predicted = fit,  cat_predicted_se = se.fit) # rename
# give each row / location a unique name to group by
data_o_plot$location_id <- paste0(data_o_plot$x, "_", data_o_plot$y)
# get difference in min max diel activity for each row
data_o_plot <- data_o_plot %>% 
  group_by(location_id) %>% 
  mutate(diel_diff = max(cat_predicted) - min(cat_predicted))

# cat plot: spatial activity per hour
o_cat_plot_hr <- ggplot(aes(x, y, fill = cat_predicted),
                     data = data_o_plot) +
  geom_tile()+
  scale_fill_viridis("log(activity)", option = "viridis", limits = c(-5.27, -1.13)) +
  geom_point(data = records_o, fill = NA, col = "white", size = 0.7, alpha = 0.1, shape = 3) +
  theme_bw(10) + 
  theme(axis.title = element_blank()) + 
    labs(title = 'Otway Ranges; feral cat; Hour: {frame_time}:00') +
  transition_time(hour) +
  ease_aes('linear')

# cat plot: spatial difference
o_cat_plot_diff <- ggplot(aes(x, y, fill = diel_diff),
                     data = data_o_plot) +
  geom_tile()+
  scale_fill_viridis("Relative diel pattern strength", option = "viridis", limits = c(0.5, 1.55), breaks=c(0.5, 1.03, 1.55),labels=c("low", "medium", "high")) +
  geom_point(data = records_o, fill = NA, col = "white", size = 0.7, alpha = 0.015, shape = 3) +
  theme_bw(10) + 
  ggtitle("d.  Otway Ranges; feral cat") +
  theme(axis.title = element_blank())

>>>>

## Spatial variation in the strength of diel patterns

@@

# predict model results into dataframe
data_g_plot <- cbind(data_g_plot, predict.gam(gam_fox_sp, newdata = data_g_plot, se.fit = TRUE, type = "response", exclude = c("s(station)", "s(survey_duration)")))
data_g_plot <- rename(data_g_plot, fox_predicted = fit,  fox_predicted_se = se.fit) # rename
# give each row / location a unique name to group by
data_g_plot$location_id <- paste0(data_g_plot$x, "_", data_g_plot$y)
# get difference in min max diel activity for each row
data_g_plot <- data_g_plot %>% 
  group_by(location_id) %>% 
  mutate(diel_diff = max(fox_predicted) - min(fox_predicted),          
         diel_diff_pc = diel_diff / min(fox_predicted) * 100)

# fox plot: spatial activity per hour
g_fox_plot_hr <- ggplot(aes(x, y, fill = fox_predicted),
                     data = data_g_plot) +
  geom_tile()+
  scale_fill_viridis("Activity", option = "viridis") +
  geom_point(data = records_g, fill = NA, col = "white", size = 0.7, alpha = 1, shape = 3) +
  theme_bw(10) + 
  theme(axis.title = element_blank()) + 
    labs(title = 'Glenelg region; red fox; Hour: {frame_time}:00') +
  transition_time(hour) +
  ease_aes('linear')

# fox plot: spatial difference
g_fox_plot_diff <- ggplot(aes(x, y, fill = diel_diff_pc),
                     data = data_g_plot) +
  geom_tile()+
  scale_fill_viridis("Relative activity increase (%)", option = "viridis", limits = c(76, 371)) +
  geom_point(data = records_g, fill = NA, col = "white", size = 0.7, alpha = 1, shape = 3) +
  theme_bw(10) + 
  ggtitle("a.  Glenelg region; red fox") +
  theme(axis.title = element_blank())

@@

# subset to just locations within buffer zone
data_o_plot = data_o_plot[with(data_o_plot, inSide(otways_buffer_df, x, y)),]
data_o_plot <- cbind(data_o_plot, predict.gam(gam_fox_sp, newdata = data_o_plot, se.fit = TRUE, type = "response", exclude = c("s(station)", "s(survey_duration)")))
data_o_plot <- rename(data_o_plot, fox_predicted = fit,  fox_predicted_se = se.fit) # rename
# give each row / location a unique name to group by
data_o_plot$location_id <- paste0(data_o_plot$x, "_", data_o_plot$y)
# get difference in min max diel activity for each row
data_o_plot <- data_o_plot %>% 
  group_by(location_id) %>% 
  mutate(diel_diff = max(fox_predicted) - min(fox_predicted),
         diel_diff_pc = diel_diff / min(fox_predicted) * 100)

# fox plot: spatial activity per hour
o_fox_plot_hr <- ggplot(aes(x, y, fill = fox_predicted),
                     data = data_o_plot) +
  geom_tile()+
  scale_fill_viridis("activity", option = "viridis") +
  geom_point(data = records_o, fill = NA, col = "white", size = 0.7, alpha = 1, shape = 3) +
  theme_bw(10) + 
  theme(axis.title = element_blank()) + 
    labs(title = 'Otway Ranges; red fox; Hour: {frame_time}:00') +
  transition_time(hour) +
  ease_aes('linear')

# fox plot: spatial difference
o_fox_plot_diff <- ggplot(aes(x, y, fill = diel_diff_pc),
                     data = data_o_plot) +
  geom_tile()+
  scale_fill_viridis("Relative activity increase (%)", option = "viridis", limits = c(76, 371)) +
  geom_point(data = records_o, fill = NA, col = "white", size = 0.7, alpha = 1, shape = 3) +
  theme_bw(10) + 
  ggtitle("c.  Otway Ranges; red fox") +
  theme(axis.title = element_blank())

#### CAT 
### GLENELG
# predict model results into dataframe
data_g_plot <- cbind(data_g_plot, predict.gam(gam_cat_sp, newdata = data_g_plot, se.fit = TRUE, type = "response", exclude = c("s(station)", "s(survey_duration)")))
data_g_plot <- rename(data_g_plot, cat_predicted = fit,  cat_predicted_se = se.fit) # rename
# give each row / location a unique name to group by
data_g_plot$location_id <- paste0(data_g_plot$x, "_", data_g_plot$y)
# get difference in min max diel activity for each row
data_g_plot <- data_g_plot %>% 
  group_by(location_id) %>% 
  mutate(diel_diff = max(cat_predicted) - min(cat_predicted),          
         diel_diff_pc = diel_diff / min(cat_predicted) * 100)

# cat plot: spatial activity per hour
g_cat_plot_hr <- ggplot(aes(x, y, fill = cat_predicted),
                     data = data_g_plot) +
  geom_tile()+
  scale_fill_viridis("Activity", option = "viridis") +
  geom_point(data = records_g, fill = NA, col = "white", size = 0.7, alpha = 1, shape = 3) +
  theme_bw(10) + 
  theme(axis.title = element_blank()) + 
    labs(title = 'Glenelg region; feral cat; Hour: {frame_time}:00') +
  transition_time(hour) +
  ease_aes('linear')

# cat plot: spatial difference
g_cat_plot_diff <- ggplot(aes(x, y, fill = diel_diff_pc),
                     data = data_g_plot) +
  geom_tile()+
  scale_fill_viridis("Relative activity increase (%)", option = "viridis", limits = c(76, 371)) +
  geom_point(data = records_g, fill = NA, col = "white", size = 0.7, alpha = 1, shape = 3) +
  theme_bw(10) + 
  ggtitle("b.  Glenelg region; feral cat") +
   theme(axis.title = element_blank())

### OTWAYS
# predict model results into dataframe
data_o_plot <- cbind(data_o_plot, predict.gam(gam_cat_sp, newdata = data_o_plot, se.fit = TRUE, type = "response", exclude = c("s(station)", "s(survey_duration)")))
data_o_plot <- rename(data_o_plot, cat_predicted = fit,  cat_predicted_se = se.fit) # rename
# give each row / location a unique name to group by
data_o_plot$location_id <- paste0(data_o_plot$x, "_", data_o_plot$y)
# get difference in min max diel activity for each row
data_o_plot <- data_o_plot %>% 
  group_by(location_id) %>% 
  mutate(diel_diff = max(cat_predicted) - min(cat_predicted),          
         diel_diff_pc = diel_diff / min(cat_predicted) * 100)

# cat plot: spatial activity per hour
o_cat_plot_hr <- ggplot(aes(x, y, fill = cat_predicted),
                     data = data_o_plot) +
  geom_tile()+
  scale_fill_viridis("Activity", option = "viridis") +
  geom_point(data = records_o, fill = NA, col = "white", size = 0.7, alpha = 1, shape = 3) +
  theme_bw(10) + 
  theme(axis.title = element_blank()) + 
    labs(title = 'Otway Ranges; feral cat; Hour: {frame_time}:00') +
  transition_time(hour) +
  ease_aes('linear')

# cat plot: spatial difference
o_cat_plot_diff <- ggplot(aes(x, y, fill = diel_diff_pc),
                     data = data_o_plot) +
  geom_tile()+
  scale_fill_viridis("Relative activity increase (%)", option = "viridis", limits = c(76, 371)) +
  geom_point(data = records_o, fill = NA, col = "white", size = 0.7, alpha = 1, shape = 3) +
  theme_bw(10) + 
  ggtitle("d.  Otway Ranges; feral cat") +
  theme(axis.title = element_blank())
